# Creating the Maps

The developable land inventory identifies land that meets the statutory definition of developable land under PA 25-1. This inventory serves as a foundational input to subsequent phases of the Housing Growth Plan process, including the suitability analysis \[link\] and the parcel and zone identification. These later steps incorporate additional planning considerations, local context, and policy objectives to evaluate where housing growth and affordable housing opportunities may be most appropriate.

To support Housing Growth Plan reporting and review, the developable land inventory should be presented in the report as municipality-wide maps. The maps should clearly display the location and distribution of developable land using clear, accessible, and consistent symbology.

The maps should incorporate the following data:

- Land identified as excluded from developable land ([Section 3](stat_def.qmd))
- Identified vacant parcels ([Section 4](vac_redev.qmd))
- Identified parcels with redevelopment potential ([Section 4](vac_redev.qmd))

The final maps included in the Housing Growth Plan will be:

1.  **Statutory Exclusions Map**: A municipality-wide map displaying each statutory exclusion category as a separate layer and symbolized distinctly to illustrate the distribution of land excluded from the developable land inventory.
2.  **Developable vs. Non-Developable Land Map**: A municipality-wide map depicting all statutory exclusion categories combined into a single consolidated layer symbolized as non-developable land. All remaining land not subject to statutory exclusions is displayed as developable land, providing a simplified binary representation of land suitability across the municipality.
3.  **Developable Land Summary Map**: A municipality-wide map displaying all statutory exclusion categories combined into a single exclusion layer, alongside a developable land layer representing all land remaining after the exclusion analysis. Identified vacant parcels and parcels with redevelopment potential should be displayed as separate overlays.

<br/>

## Map Layers and Symbology

All maps include the municipal boundary layer to delineate the geographic extent of the study area.

**Dataset**: [CT Municipalities (CT Geodata Portal)](https://geodata.ct.gov/datasets/ctmaps::ct-municipalities-with-fips/about)

<br/>

Municipal parcel boundaries are included where appropriate to provide spatial context for interpretation of mapped datasets.

**Dataset**: [Connecticut CAMA and Parcel Layer (CT Geodata Portal)](https://geodata.ct.gov/datasets/ctmaps::connecticut-cama-and-parcel-layer/about)

<br/>

1.  **Statutory Exclusions Map**

```{r}
#| echo: false
#| warning: false
#| message: false

library(knitr)
library(kableExtra)

map_layers <- data.frame(
  `Map Layer (top to bottom)` = c(
    "Municipality boundary",
    "Parcels",
    "Land already committed to public use or purpose",
    "Open space, parks, and recreation areas dedicated to the public or protected by a conservation easement",
    "Land subject to enforceable restrictions or prohibitions on development",
    "Wetlands or watercourses, as defined in the Wetlands and Watercourses Act (CGS Chapter 440)",
    "Areas of 0.5 acres or more of contiguous land that are unsuitable for development due to topography (such as steep slopes)"
  ),
  `Description` = c(
    "Municipal boundary delineating the extent of the study area.",
    "Municipal parcel boundaries.",
    "Exclusion category described in Section 3.1.",
    "Exclusion category described in Section 3.2.",
    "Exclusion category described in Section 3.3.",
    "Exclusion category described in Section 3.4.",
    "Exclusion category described in Section 3.5."
  ),
  `Recommended Symbology` = c(
    "",
    "No fill; thin light-gray outlines.",
    "",
    "",
    "",
    "",
    ""
  ),
  check.names = FALSE
)

kable(
  map_layers,
  align = "ccc",
  col.names = c(
    "Map Layer (top to bottom)",
    "Description",
    "Recommended Symbology"
  )
) |>
  kable_styling(full_width = TRUE) |>
  column_spec(1, width = "33%") |>
  column_spec(2, width = "33%") |>
  column_spec(3, width = "34%")
```

<br/>

2.  **Developable vs. Non-Developable Land Map**
```{r}
#| echo: false
#| warning: false
#| message: false

library(knitr)
library(kableExtra)

developable_layers <- data.frame(
  `Map Layer (top to bottom)` = c(
    "Municipality boundary",
    "Statutory exclusions",
    "Developable land"
  ),
  `Description` = c(
    "Municipal boundary delineating the extent of the study area.",
    "Areas excluded from the developable land inventory in Section 2. All statutory exclusion categories should be combined into a single layer.",
    "Land not identified as excluded under the statutory criteria. This layer represents all land remaining after the exclusion analysis."
  ),
  `Recommended Symbology` = c(
    "",
    "Dark gray ([hex code])",
    "Light gray"
  ),
  check.names = FALSE
)

kable(
  developable_layers,
  col.names = c(
    "Map Layer (top to bottom)",
    "Description",
    "Recommended Symbology"
  ),
  align = "ccc"
) |>
  kable_styling(full_width = TRUE) |>
  column_spec(1, width = "33%") |>
  column_spec(2, width = "33%") |>
  column_spec(3, width = "34%")
```

<br/>

3.  **Developable Land Summary Map**
```{r}
#| echo: false
#| warning: false
#| message: false

library(knitr)
library(kableExtra)

vacant_redevelopment_layers <- data.frame(
  `Map Layer (top to bottom)` = c(
    "Municipality boundary",
    "Parcels",
    "Vacant Areas",
    "Areas with Redevelopment Potential",
    "Statutory Exclusions",
    "Developable Land"
  ),
  `Description` = c(
    "Municipal boundary delineating the extent of the study area.",
    "Municipal parcel boundaries.",
    "Areas identified as vacant in Section 4.1.",
    "Areas identified as having redevelopment potential in Section 4.2.",
    "Areas excluded from the developable land inventory in Section 2. All statutory exclusion categories should be combined into a single layer.",
    "Land not identified as excluded under the statutory criteria. This layer represents all land remaining after the exclusion analysis."
  ),
  `Recommended Symbology` = c(
    "",
    "No fill; thin light-gray outlines.",
    "",
    "",
    "Gray ([hex code])",
    ""
  ),
  check.names = FALSE
)

kable(
  vacant_redevelopment_layers,
  col.names = c(
    "Map Layer (top to bottom)",
    "Description",
    "Recommended Symbology"
  ),
  # align = "ccc"
) |>
  kable_styling(full_width = TRUE) |>
  column_spec(1, width = "33%") |>
  column_spec(2, width = "33%") |>
  column_spec(3, width = "34%")
```


::: {.callout-caution appearance="simple" icon="false"}
Map symbology should use colors and patterns that improve overall readability, provide clear visual distinction between categories, and remain accessible to users with color vision deficiencies.
:::

<br/>

## Required Map Elements

All maps should include:

- Municipality-wide extent
- Clear map title
- Scale bar or representative fraction
- North arrow
- Legend identifying all mapped categories
- Data sources and preparation date (recommended)
